Notes on the low density parameters in {\tt MAESTRO}

### 2011-03-22 docs/lo_density/lo_density.tex
first attempt at describing new definition of cutoff densities for multilevel planar

\section{Computing the Cutoff Values}

At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}(:), 
{\tt base\_cutoff\_density\_coord}(:), and {\tt burning\_cutoff\_density\_coord}(:) by 
setting the corresponding coordinate equal to $r$ as soon as $\rho_0(r)$ is less than 
or equal to that particular cutoff value (see Figure \ref{Fig:Cutoff} for the
single-level case).  Multilevel is more complicated.

We set the coordinates at the following points in the code:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hpb]
\centering
\includegraphics[width=4.0in]{\lodensfigpath/cutoff}\hspace{0.2in}
\begin{minipage}[b]{5.0in}
\caption[Relation between cutoff density and cutoff coordinates]
{Image of how the cutoff density and cutoff coordinates
are related.  This is drawn assuming {\tt drdxfac}=5.\vspace{2em}}
\end{minipage}
\label{Fig:Cutoff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Density vs. Coordinate}

Note that saying $r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)\le$ {\tt anelastic\_cutoff}.  Also, saying $r<$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt anelastic\_cutoff}.  Ditto for {\tt base\_cutoff\_density} and {\tt base\_cutoff\_density\_coord}.

\section{anelastic\_cutoff}\label{Sec:Anelastic Cutoff}

@@

\section{base\_cutoff\_density}\label{Sec:Base Cutoff Density}

@@

\section{burning\_cutoff}

@@

\section{buoyancy\_cutoff\_factor}

>>>>

\section{Computing the Cutoff Values}
We compute {\tt anelastic\_cutoff\_coord}(:), {\tt base\_cutoff\_density\_coord}(:), 
and {\tt burning\_cutoff\_density\_coord}(:) in analogous fashion.

\subsection{Single-Level Planar or any Spherical}
Here the base state exists as a single one-dimensional array with constant grid
spacing $\Delta r$.  Basically, we set the corresponding coordinate equal to $r$ as soon 
as $\rho_0(r)$ is less than  or equal to that particular cutoff value.
See Figure \ref{Fig:Cutoff} for a graphical representation.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hpb]
\centering
\includegraphics[width=4.0in]{\lodensfigpath/cutoff}\hspace{0.2in}
\caption{Image of how the cutoff density and cutoff coordinates
are related for single-level planar and all spherical problems.}
\label{Fig:Cutoff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Note that for single-level planar or any spherical problem, saying $r\ge$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)\le$ {\tt anelastic\_cutoff}.  Also, saying $r<$ {\tt anelastic\_cutoff\_coord} is analogous to saying $\rho_0(r)>$ {\tt anelastic\_cutoff}.  Ditto for {\tt base\_cutoff\_density} and {\tt base\_cutoff\_density\_coord}.

\subsection{Multilevel Planar}
In this case, the base state exists as several one-dimensional arrays, each with
different grid spacing.  Refer to Figure \ref{Fig:Cutoff_Multi} in the following examples.
The guiding principle is to check whether $\rho_0$ falls below $\rho_{\rm cutoff}$ on the finest
grid first.  If not, check the next coarser level.  Continue until you reach the base grid,
in which case $\rho_0$ must fall below $\rho_{\rm cutoff}$ somewhere.  Some examples are in order:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hpb]
\centering
\includegraphics[width=2.0in]{\lodensfigpath/cutoff_multi}\hspace{0.2in}
\caption{Multilevel cutoff density example.}
\label{Fig:Cutoff_Multi}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\item {\bf Example 1:} $\rho_{0,104} > \rho_{\rm cutoff}$ and $\rho_{0,105} < \rho_{\rm cutoff}$.\\ \\
{\tt cutoff\_density\_coord}(1) = 105\\
{\tt cutoff\_density\_coord}(2) = 210\\
{\tt cutoff\_density\_coord}(3) = 420\\ \\
This is the simplest case in which the cutoff transition happens on the coarsest level.
In this case, the cutoff coordinates at the finer levels are simply propagated from the
coarsest level, even though they do not correspond to a valid region.

\item {\bf Example 2:} $\rho_{0,403} > \rho_{\rm cutoff}$ and $\rho_{0,404} < \rho_{\rm cutoff}$.\\ \\
{\tt cutoff\_density\_coord}(1) = 101\\
{\tt cutoff\_density\_coord}(2) = 202\\
{\tt cutoff\_density\_coord}(3) = 404\\ \\
In this case, the cutoff transition happens where the finest grid is present.  Happily, the
transition occurs at a location where there is a common grid boundary between all three levels.
Therefore, we simply propagate the cutoff density coordinate from the finest level downward.

\item {\bf Example 3:} $\rho_{0,404} > \rho_{\rm cutoff}$ and $\rho_{0,405} < \rho_{\rm cutoff}$.\\ \\
{\tt cutoff\_density\_coord}(1) = 102\\
{\tt cutoff\_density\_coord}(2) = 203\\
{\tt cutoff\_density\_coord}(3) = 405\\ \\
In this case, the cutoff transition happens where the finest grid is present.  However, the
transition occurs at a location where there NOT is a common grid boundary between all three 
levels.  We choose to define the cutoff transition at the coarser levels as being at the
corresponding boundary that is at a larger radius than the location on the finest grid.

\section{When are the Cutoff Coordinates Updated?}

At several points in the algorithm, we compute {\tt anelastic\_cutoff\_coord}(:), 
{\tt base\_cutoff\_density\_coord}(:), and {\tt burning\_cutoff\_density\_coord}(:):

@@

\section{Usage of Cutoff Densities}

\subsection{anelastic\_cutoff}\label{Sec:Anelastic Cutoff}

@@

\subsection{base\_cutoff\_density}\label{Sec:Base Cutoff Density}

@@

\subsection{burning\_cutoff}

@@

\subsection{buoyancy\_cutoff\_factor}
